---
title: "Converting qPCR raw data into data analysis"
output: html_notebook
---



Exported file from the quantstudio 5 is of the following format:
Various metrics are located in different sheets with specific names, i.e. "Results"
The Results page contains rows of information at the start irrelevant to the data
A row containing all the collected variables and measurements is the start of the "data matrix"
the column names include "Well", "Well Position" etc.

The goal is to write a file that can take an exported excel file, read in the data and do the analysis to produce
measurements like Normalised expression (DCt) and Relative expression if needed (DDCt)

Even if this simply imports the data and renders it usable for graphing that would be great, but if i can get
it to graph as well that would be fantastic


Step 1: Read the .xlsx file
```{r}

library(here)
library(readxl)
df <- read_excel(here("rstudio-git-test/QP069.xls"), sheet = "Results",
                 skip = 46, n_max = 96)

```

We only need certain rows, so will remove the others
```{r}
df <- df[,c("Well Position", "Sample Name", "Target Name", "CT", "Ct Mean", "Ct SD")]
df
```

Make a df that contains only the mean values for each combination
Try to clean up the data (remove probably failures)
```{r}
df.sum <- distinct(df[,-c(1,4)])
# will remove all entries where the sd of Ct is >0.9, which is a good indication
# that the technical replicates are not accurate
for (i in 1:nrow(df.sum)) {
  if (df.sum[i,4]>0.9) {
    print(df.sum[i,])
    df.sum[i,] <- NA
  } 
}
df.sum <- df.sum[!is.na(df.sum[,1]),]

```

Break up into sample (in this case target as i enterred the names incorrectly at the interface)
Calculate DDct 
Might try to flip this around, if initial attempt doesnt work
```{r}
library(dplyr)

# iterates through the df and adds a tibble df of each sample to a list
sum.list <- as.list(NULL)
temp <- as.data.frame(NULL)
for (i in 1:length(samples)) {
  
  for(j in 1:nrow(df.sum)) {
    if (df.sum[j, 2] == samples[i]) {
      temp <- rbind(temp, df.sum[j,])
     j <- j+1
    }
    
  sum.list[[samples[i]]] <- temp
  }
  temp <- as.data.frame(NULL)
  i <- i+1
}
```

Now have a list containing individual dfs seperated by Sample
To generate DCt need : test Ct - control Ct, which leads to the normalised expression of a gene
normalised expression (NormExp) = 2^DCt
in this case (and almost all cases) control is GAPDH and will be the first row in the list.

```{r}
## at this point, can ask for input as to what the control gene is (drop down list of genes used)
sum.list.dct <- sum.list
for (i in 1:length(sum.list)) {

  for (j in 1:nrow(sum.list[[i]])) {
    
    sum.list.dct[[i]][j,"DCT"] <- sum.list[[i]][j,3] - sum.list[[i]][sum.list[[i]][,1]=="GAPDH",3]
    sum.list.dct[[i]][j,"NormExp"] <- 2^-(sum.list.dct[[i]][j,"DCT"])
    print(sum.list.dct[[i]][j,"NormExp"])
  }
}
```

Repeat this with DDCt which is test DCT - control DCt (RelExp)
Produce produce drop down menu OR have a variable input signifying which sample is control/relative

```{r}

sum.list.ddct <- lapply(sum.list.dct, function(x) x[x$`Sample Name` != "GAPDH",])

for (i in 1:length(sum.list.ddct)) {
  
  for (j in 1:nrow(sum.list.dct[[i]])) {
    
    
  }
  
  sum.list.ddct[[i]][,"DDCT"] <- sum.list.ddct[[i]][,"DCT"] - sum.list.ddct[[1]][,"DCT"]
  
  print(sum.list.ddct[[i]][,"DDCT"])
}


```

Attempt to combine into single data frame and then split into a list of Genes

```{r}

```








