---
title: "Converting qPCR raw data into data analysis"
output: html_notebook
---

HEADING
===================
This file converts an exported excel file from the QS5 and manipulates it to:
* Remove reads where technical replicates were too different
* Remove samples where the housekeeping gene (GAPDH until I modify to make a function) is too lowly expressed

What this file does
===================

Exported file from the quantstudio 5 is of the following format:
Various metrics are located in different sheets with specific names, i.e. "Results"
The Results page contains rows of information at the start irrelevant to the data
A row containing all the collected variables and measurements is the start of the "data matrix"
the column names include "Well", "Well Position" etc.

The goal is to write a file that can take an exported excel file, read in the data and do the analysis to produce
measurements like Normalised expression (DCt) and Relative expression if needed (DDCt)

Even if this simply imports the data and renders it usable for graphing that would be great, but if i can get
it to graph as well that would be fantastic


Programming:
===========

Step 1: Read the .xlsx file
```{r}

library(here)
library(readxl)
df <- read_excel(here("rstudio-git-test/QP069.xls"), sheet = "Results",
                 skip = 46, n_max = 96)

```

We only need certain rows, so will remove the others
```{r}
df <- df[,c("Well Position", "Sample Name", "Target Name", "CT", "Ct Mean", "Ct SD")]
df
```

Make a df that contains only the mean values for each combination
Try to clean up the data (remove probably failures)
```{r}
df.sum <- distinct(df[,-c(1,4)])
# will remove all entries where the sd of Ct is >0.9, which is a good indication
# that the technical replicates are not accurate
for (i in 1:nrow(df.sum)) {
  if (df.sum[i,4]>0.9) {
    print(df.sum[i,])
    df.sum[i,] <- NA
  } 
}
<<<<<<< HEAD

df.sum <- df.sum[!is.na(df.sum[,1]),]


```

May try to remove those where GAPDH is too high:
```{r}
weird <- as.vector(NULL)
for(i in 1:nrow(df.sum)) {
  if (df.sum[i,1] == "GAPDH" & df.sum[i,3] > 22) {
    print(df.sum[i,2])
    weird <- c(weird, as.character(df.sum[i,2]))
  }
}
for (i in 1:nrow(df.sum)) {
  for(j in length(weird)) {
    if (df.sum[i,2] == weird[j]){
      print(df.sum[i,])
      df.sum[i,] <- NA
    }
  } 

}

df.sum <- df.sum[!is.na(df.sum[,1]),]
```








Break up into sample (in this case target as i enterred the names incorrectly at the interface)
Calculate DDct 
Might try to flip this around, if initial attempt doesnt work
```{r}
library(dplyr)

samples <- as.vector(df.sum %>% 
  group_by(`Target Name`) %>% 
  summarise(n_distinct(`Sample Name`)))$`Target Name`


# iterates through the df and adds a tibble df of each sample to a list
sum.list <- as.list(NULL)
temp <- as.data.frame(NULL)
for (i in 1:length(samples)) {
  
  for(j in 1:nrow(df.sum)) {
    if (df.sum[j, 2] == samples[i]) {
      temp <- rbind(temp, df.sum[j,])
     
    }
    
  sum.list[[samples[i]]] <- temp
  }
  temp <- as.data.frame(NULL)
  
}
```

Now have a list containing individual dfs seperated by Sample
To generate DCt need : test Ct - control Ct, which leads to the normalised expression of a gene
normalised expression (NormExp) = 2^DCt
in this case (and almost all cases) control is GAPDH and will be the first row in the list.

```{r}
## at this point, can ask for input as to what the control gene is (drop down list of genes used)
sum.list.dct <- sum.list
for (i in 1:length(sum.list)) {

  for (j in 1:nrow(sum.list[[i]])) {
    
    sum.list.dct[[i]][j,"DCT"] <- sum.list[[i]][j,3] - sum.list[[i]][sum.list[[i]][,1]=="GAPDH",3]
    sum.list.dct[[i]][j,"NormExp"] <- 2^-(sum.list.dct[[i]][j,"DCT"])
    print(sum.list.dct[[i]][j,"NormExp"])
  }
}
```

Repeat this with DDCt which is test DCT - control DCt (RelExp)
Produce produce drop down menu OR have a variable input signifying which sample is control/relative
Attempt to combine into single data frame and then split into a list of Genes
```{r}
# removes control (GAPDH) from dataset for further analysis
sum.list.dct <- lapply(sum.list.dct, function(x) x[x$`Sample Name` != "GAPDH",])

#generates df containing all current information
df.dct <- as.data.frame(NULL)
for (i in 1:length(sum.list.dct)) {
  for (j in 1:nrow(sum.list.dct[[i]])) {
    df.dct <- rbind(df.dct, sum.list.dct[[i]][j,])
  }
}
# generates vector of genes to use for matching
genes <- as.vector(df.dct %>% 
  group_by(`Sample Name`) %>% 
  summarise(n_distinct(`Target Name`)))$`Sample Name`

# produces list containing dfs for each gene
sum.list.dct <- as.list(NULL)
temp <- as.data.frame(NULL)
for (i in 1:length(genes)) {
  
  for(j in 1:nrow(df.dct)) {
    if (df.dct[j, 1] == genes[i]) {
      temp <- rbind(temp, df.dct[j,])
     
    }
    
  sum.list.dct[[genes[i]]] <- temp
  }
  temp <- as.data.frame(NULL)
  
}

=======
>>>>>>> 8f63bfbededfa866bdf54a0dc2390238d3ffd9db

# iterates through the df and adds a tibble df of each sample to a list
sum.list <- as.list(NULL)
temp <- as.data.frame(NULL)
for (i in 1:length(samples)) {
  
  for(j in 1:nrow(df.sum)) {
    if (df.sum[j, 2] == samples[i]) {
      temp <- rbind(temp, df.sum[j,])
     j <- j+1
    }
    
  sum.list[[samples[i]]] <- temp
  }
  temp <- as.data.frame(NULL)
  i <- i+1
}
```

Now have a list containing individual dfs seperated by Sample
To generate DCt need : test Ct - control Ct, which leads to the normalised expression of a gene
normalised expression (NormExp) = 2^DCt
in this case (and almost all cases) control is GAPDH and will be the first row in the list.

```{r}
## at this point, can ask for input as to what the control gene is (drop down list of genes used)
sum.list.dct <- sum.list
for (i in 1:length(sum.list)) {

  for (j in 1:nrow(sum.list[[i]])) {
    
    sum.list.dct[[i]][j,"DCT"] <- sum.list[[i]][j,3] - sum.list[[i]][sum.list[[i]][,1]=="GAPDH",3]
    sum.list.dct[[i]][j,"NormExp"] <- 2^-(sum.list.dct[[i]][j,"DCT"])
    print(sum.list.dct[[i]][j,"NormExp"])
  }
}
```

Repeat this with DDCt which is test DCT - control DCt (RelExp)
Produce produce drop down menu OR have a variable input signifying which sample is control/relative

```{r}

sum.list.ddct <- lapply(sum.list.dct, function(x) x[x$`Sample Name` != "GAPDH",])

for (i in 1:length(sum.list.ddct)) {
  
  for (j in 1:nrow(sum.list.dct[[i]])) {
    
    
  }
  
  sum.list.ddct[[i]][,"DDCT"] <- sum.list.ddct[[i]][,"DCT"] - sum.list.ddct[[1]][,"DCT"]
  
  print(sum.list.ddct[[i]][,"DDCT"])
}


```

Attempt to combine into single data frame and then split into a list of Genes

```{r}

```





Next is to generate the relative expression column (RelExp = testDct - controlDct)
Currently this ignores the possibility of biological replicates
```{r}
sum.list.ddct <- sum.list.dct
for (i in 1:length(sum.list.dct)) {

  for (j in 1:nrow(sum.list.dct[[i]])) {
    
    sum.list.ddct[[i]][j,"DDCT"] <- sum.list.dct[[i]][j,5] - sum.list.dct[[i]][sum.list.dct[[i]][,2]=="HOM 10-2",5]
    
    sum.list.ddct[[i]][j,"RelExp"] <- 2^-(sum.list.ddct[[i]][j,"DDCT"])
    print(sum.list.ddct[[i]][j,"RelExp"])
  }
}

df.ddct <- as.data.frame(NULL)
for (i in 1:length(sum.list.ddct)) {
  for (j in 1:nrow(sum.list.ddct[[i]])) {
    df.ddct <- rbind(df.ddct, sum.list.ddct[[i]][j,])
  }
}
```


Graph relative expression 

```{r}
library(ggplot2)

ggplot(df.ddct, aes(`Target Name`, RelExp, fill = `Sample Name`)) +
  geom_col() +
  facet_grid(`Sample Name` ~ ., scales = "free_y")

ggplot(df.ddct, aes(`Target Name`, NormExp, fill = `Sample Name`)) +
  geom_col() +
  facet_grid(`Sample Name` ~ ., scales = "free_y")

```






