---
title: "Converting qPCR raw data into data analysis"
output: html_notebook
---



Exported file from the quantstudio 5 is of the following format:
Various metrics are located in different sheets with specific names, i.e. "Results"
The Results page contains rows of information at the start irrelevant to the data
A row containing all the collected variables and measurements is the start of the "data matrix"
the column names include "Well", "Well Position" etc.

The goal is to write a file that can take an exported excel file, read in the data and do the analysis to produce
measurements like Normalised expression (DCt) and Relative expression if needed (DDCt)

Even if this simply imports the data and renders it usable for graphing that would be great, but if i can get
it to graph as well that would be fantastic


Step 1: Read the .xlsx file
```{r}
install.packages("here")
library(here)
library(readxl)
df <- read_excel(here("rstudio-git-test/QP069.xls"), sheet = "Results",
                 skip = 46, n_max = 96)

```

We only need certain rows, so will remove the others
```{r}
df <- df[,c("Well Position", "Sample Name", "Target Name", "CT", "Ct Mean", "Ct SD")]

```





